\documentclass[11pt,a4paper,english]{paper}
\usepackage{mathtools}
\usepackage[breakable]{tcolorbox}
%\usepackage{minted}
\newtcolorbox{mybox}[1]{colback=red!5!white,colframe=red!75!black,fonttitle=\bfseries,title=#1,breakable}
\newtcolorbox{bluebox}[1]{colback=blue!5!white,colframe=blue!75!black,fonttitle=\bfseries,title=#1,breakable}
\newtcolorbox{gbox}[1]{colback=green!5!white,colframe=green!75!black,fonttitle=\bfseries,title=#1,breakable}
\newtcolorbox{bbox}[1]{colback=black!5!white,colframe=black!75!black,fonttitle=\bfseries,title=#1,breakable}
\usepackage{amsmath}                                    % extensive math options
\usepackage{amssymb}                                    % special math symbols
\usepackage[mathlines]{lineno}
\usepackage[Gray,squaren,thinqspace,thinspace]{SIunits} % elegant units
\usepackage{listings}                                   % source code

\usepackage{graphicx}
\graphicspath{ {./} }
%\setminted{breaklines}

\begin{document}

\title{CS 008 \\ Lecture notes \\ 4/25/24}
\maketitle

\begin{gbox}{Pre-lecture notes:} {

Exam scores are out, grades are released on canvas.

Most of the grading will be based off of assignments and lab-work.


}\end{gbox}

\section{Direct Access Arrays and Hash Tables pt. 2}

\begin{bbox}{Outline:}{

Another lecture on these topics:

\begin{itemize}

  \item Motivation
  \item Word RAM model
  \item Direct access arrays
  \item Hash tables

\end{itemize}
} \end{bbox}

\section{Motivation}

\section{Word RAM model}

\section{Direct access arrays}

A direct access array is an array where every index in the array is associated with a specific element. Each element has it's own place in the array. Direct access arrays are very fast but have a major flaw where it takes up too much space.

\bigskip
\begin{bluebox}{Operations:} {

    As previously stated, direct access arrays are fast however they are very spacially inefficient.

    \begin{itemize}

    \item $build() = O(n)$
    \item $find(k) = O(1)$
    \item $add(k) = O(1)$
    \item $delete(k) = O(1)$

    \end{itemize}


}\end{bluebox}

\bigskip
\noindent
In summary, direct access arrays are great for speed however for larger ranges of values, the size of the array would be too big. With smaller ranges of numbers.

\section{Hash tables}



\end{document}
